### 2024-06-05 Scripts/Manuscript_scripts.Rmd
Update Manuscript_scripts.Rmd

>>>>

df_profiles = read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Main_figures\microbiota-pneumonia-species_phyloseqized.csv)",header = FALSE)
sampleid2site = read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Main_figures\metadata-pneumonia-sampleid2site.csv)",header = TRUE)
df_arrows <- read.csv(r'(LRT_microbiomes_critical_illness\Data_repository\Main_figures\MDS_metrics-bray-specie_pneumonia-arrows_coordinates.csv)')

#otu-like data reformat 
otu_metaphlan <- as.matrix(df_profiles[2:nrow(df_profiles),2:ncol(df_profiles)]) # otu matrix can't have any character-like data, you have to use index to remove it
class(otu_metaphlan) <- "numeric" # convert the datatype of the matrix into "numeric"
otu_metaphlan <- otu_metaphlan*1000000
rownames(otu_metaphlan) <- df_profiles[2:nrow(df_profiles),1]
colnames(otu_metaphlan) <- df_profiles[1,2:ncol(df_profiles)]

OTU = otu_table(otu_metaphlan,taxa_are_rows = TRUE)
rownames(sampleid2site) <- sampleid2site$SampleID # you have to re-assign the rownames of meta-data, or you will run into a problem: "invalid “phyloseq” object: Component sample names do not match"
sam_data = sample_data(sampleid2site)
physeq_profiles = phyloseq(OTU,sam_data) 
target_dist_methods <- c('bray')
for( i in target_dist_methods ){
  target_dist <- distance(physeq_profiles, method=i)
  target_MDS_metrics <- ordinate(physeq_profiles, "MDS", distance=target_dist)
  df_metrics <- data.frame(target_MDS_metrics[["vectors"]])
  df_metrics_filter <- df_metrics[,1:2]

  # Merge df_metrics_filter and sampleid2site on the index
  df_metrics_filter <- merge(df_metrics_filter, sampleid2site, by="row.names", all.x=TRUE)
  df_metrics_filter  <- df_metrics_filter[,c(4,2,3,5)]
}

df_arrows$variable <- str_split(df_arrows$variable, "__", simplify = TRUE)[,2]
df_arrows$variable  <- str_replace_all(df_arrows$variable, "_", " ")
df_arrows$variable  <- str_replace_all(df_arrows$variable, " group", " ")
df_distance_bray$Axis.2 <- df_distance_bray$Axis.2*-1
df_arrows$Axis.2 <- df_arrows$Axis.2*-1

plots <- ggscatter(df_distance_bray,
                   x = "Axis.1", 
                   y = "Axis.2",
                   color = "SamplingSite",
                   #shape = "SamplingSite",
                   size=6,
                   # palette = c( "#FC4E07","#00AFBB", "#E7B800"), #color1
                   palette = c("#e64833", "#01a188", "#3e5386"), #color2
                   # palette = c("#c23726", "#e3e3e3", "#0036fc"),  #palette2
                   # palette = c( "#010101","#fe0000", "#0037fe"), #palette4
                   #palette = c("#c23726", "#5a5a5a", "#2c83c6"),
                   #palette = c("#7152a9", "#ffc376", "#eb7600"),
                   margin.plot = "boxplot",
                   ggtheme = theme_pubr(),
                   #margin.params = list(fill="SamplingSite",size=0.36),
                   legend = 'right') + 
  geom_segment(data = df_arrows,
               x = 0, y = 0, alpha = 0.7,
               mapping = aes(xend = Axis.1, yend = Axis.2),
               color = 'black',size=0.1,
               # Add arrow head
               arrow = arrow(length = unit(3, "mm"))) +
  geom_label(data = df_arrows, aes(label = variable),color='black',alpha=0.7,size=7,label.size =0.01) 
ggpar(plots,
      xlab = "MDS1 (27.6%)",
      ylab = "MDS2 (16.6%)") + theme_pubr(base_size = 24,border = T,legend = 'none')

```{r Fig 2b}

df_distance_bray_species_pneumonia <- read.csv(r'(LRT_microbiomes_critical_illness\Data_repository\Main_figures\MDS_metrics-bray+species_abundance-species_pneumonia.csv)')
df_distance_bray_species_pneumonia$variable <- str_split(df_distance_bray_species_pneumonia$variable, "__", simplify = TRUE)[,2]
df_distance_bray_species_pneumonia$variable  <- str_replace_all(df_distance_bray_species_pneumonia$variable, "_", " ")
df_distance_bray_species_pneumonia$variable  <- str_replace_all(df_distance_bray_species_pneumonia$variable, " group", " ")
df_distance_bray_species_pneumonia$Axis.2 <- df_distance_bray_species_pneumonia$Axis.2*-1

plots_mds_species <- ggscatter(df_distance_bray_species_pneumonia,
                   x = "Axis.1", 
                   y = "Axis.2",
                   # color = "Streptococcus",
                   color = 'value',
                   #shape = "SamplingSite",
                   size=6,
                   margin.plot = "boxplot",
                   ggtheme = theme_pubr(border = T),
                   #margin.params = list(fill="Acinetobacter",size=0.36),
                   legend = 'right') + #gradient_color(inferno(10))
                   gradient_color(c('#bfbdbd','#fe1608')) + facet_wrap(~ variable) +
  labs(x='MDS1 (19.8%)',y='MDS2 (14.5%)',color='Relative\nabundance (%)') + 
  theme(axis.text.x  = element_text(size=32),axis.text.y = element_text(size=32),
        axis.title.x = element_text(size = 32),axis.title.y = element_text(size = 32)) +
  theme(strip.background = element_blank(),
        strip.text = element_textbox(
          size = 30,
          color = "black", fill = "#d9d9d9", box.color = "#4A618C",#face = 'italic',
          halign = 0.5, linetype = 1,  width = unit(1, "npc"),
          margin = margin(0, 0, 0, 0),padding = margin(6, 6, 6, 6)),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size=14),
        legend.text = element_text(size=16)) +
  xlim(-0.5,0.53)
plots_mds_species

```{r Fig 2c,d}

patientids <- c('P68','P23')
df_abundance <- read.csv(paste("LRT_microbiomes_critical_illness\Data_repository\Main_figures/microbiota-individual_patient/",
                               patientid,"-metaphlan3-species_microbiota.csv",sep=""))

df_abundance$Sampling_Day_ICU <- factor(df_abundance$Sampling_Day_ICU,levels = unique(df_abundance$Sampling_Day_ICU))
plot_df <- ggplot(df_abundance, aes(fill=taxonomy, x=Sampling_Day_ICU, y=relative_abundance))+
  geom_bar(position = "stack",stat = "identity",) + ggtitle(label = patientid ) + 
  scale_fill_manual(values = species2color)
ggpar(plot_df,
      ylab = "Relative\nabundance (%)",
      xlab = F) + theme_pubr(base_size = 14,x.text.angle =50,legend = "none",margin = FALSE)  + xlab(NULL) +
  ggeasy::easy_center_title()

```{r Fig 2e}

### visualized with circos webserver

```{r Fig 2g}

my_comparisons <- list(c("Yes", "No"))
df_combined$tetracycline__tetA_log10 <- log10(df_combined$tetracycline__tetA)
ggboxplot(df_combined, x="Diagnosed.Pneumonia", y="tetracycline__tetA_log10",
          color = "Diagnosed.Pneumonia", 
          add = "jitter",add.params = list(size = 2.4, alpha = 0.8))  +
  geom_boxplot(aes(color=Diagnosed.Pneumonia), lwd = 0.7) +
  scale_x_discrete(labels = c("No" = "No Pneumonia", "Yes" = "Pneumonia")) +
  stat_compare_means(comparisons = my_comparisons,label = "p.label",size=7,
                     vjust=.15,tip.length= c(0,0,0)) + 
  labs(x=NULL,y='Total ARG abundance (log10)',) +  ggtitle('Tetracycline__tetA') +
  # scale_y_continuous(limits=c(-2,2))+
  theme_pubr(base_size = 23,legend ='none',x.text.angle = 30) + 
  ggeasy::easy_center_title() + 
  scale_color_npg()

```{r Fig 2h}

my_comparisons <- list(c("Yes", "No"))

df_combined$multidrug__mexB_log10 <- log10(df_combined$multidrug__mexB)
ggboxplot(df_combined, x="Diagnosed.Pneumonia", y="multidrug__mexB_log10",
          color = "Diagnosed.Pneumonia", 
          add = "jitter",add.params = list(size = 2.4, alpha = 0.8))  +
  geom_boxplot(aes(color=Diagnosed.Pneumonia), lwd = 0.7) +
  scale_x_discrete(labels = c("No" = "No Pneumonia", "Yes" = "Pneumonia")) +
  stat_compare_means(comparisons = my_comparisons,label = "p.label",size=7,
                     vjust=.15,tip.length= c(0,0,0)) + 
  labs(x=NULL,y='Total ARG abundance (log10)',) +  ggtitle('Multidrug__mexB') +
  theme_pubr(base_size = 23,legend ='none',x.text.angle = 30) + 
  ggeasy::easy_center_title() + 
  scale_color_npg()

```{r Fig 2i}
